Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}

### 2013-03-30 W13report.tex
Making sure the WCS sections are self-consistent.

We found a severe dependence of the false positive rate on our
understanding of the variance; even a 1\% misestimate of the noise may
increase or decrease the false positive rate by a factor of {\bf XX}
at 5--sigma.  Prefiltering is preferred in this regard, as our
variance planes are shown to track the variance to higher fidelity at
the detection stage.  We find that over/undersubtraction of the image
background leads to a bias in the positive to negative false detection
ratio, but not on the overall numbers of detections.  Prefiltering
also prevents, in many cases, a deconvolution of the template image
when the science image has better seeing, which is a clear advantage
of the technique.

@@

\subsection{Astrometric Registration \label{subsec-astrom}}
The original plan for this production was to use the single deep exposures (v6866600) as 
inputs to the coadd process, to exercise use of the coadd repository.  As 
such we warped the exposures to coadd patches for use in the image difference
task.  The register task was unable to solve a differential WCS between the coadd
and science {\tt calexp} because in coadds, the parent WCS is for the entire {\tt tract}.
Each coadd {\tt patch} thus has a CRPIX at the center of the {\tt tract}, which is
typically off the {\tt patch} image.  TAN--SIP is unable to perform in this circumstance,
as it expands its polynomials around the CRPIX.  To get around this 
we tried re--solving the coadd patches using a TAN--SIP solution that was solved from
scratch.  This worked fairly will part of the time,
yielding residuals in the 40--100 mas range, compared to the 250 mas previously.  However,
as we did not understand the origin of these problems, and there were legitimate bugs
in meas\_astrom  that were uncovered as a result of this investigation, we reverted to
using the deep {\tt calexp} as the template directly; the direct deep {\tt calexp} to science {\tt calexp} 
registration typically yielded residuals in the range of 4--5 mas.

We investigated potential causes of the large residuals.  To do this, we looked at the 
residuals for template--to--refcat solutions, science image--to--ref solutions, and 
direct template to science image registration to attempt to isolate where the astrometry 
was being degraded.  The results are shown
in table \ref{tab-wcsrms}.  We show the results for the nine sensors in the best seeing visit.

The solutions for the three steps that do not involve any warping show results consistent with 
the accuracy we expect to get.  We also see the expected trend that the deep exposures have 
smaller RMS than the solutions of the science images.  The coadd--to--science registration
depends critically on the WCS fitting and warping process: the WCS of the deep {\tt calexp} is warped
into the coadd {\tt patch}; the coadd {\tt patch} is then warped into the science {\tt calexp} frame.
This places two dependencies on the warping code, and two on the WCS fitting, where
this degradation may have occurred.

@@

\clearpage
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{WCS RMS in mas at various stages} \\
\hline
Sensor    & Deep & Science &  Deep to Science & Coadd to Science \\
\hline
R22:S00&4.2&5.8&3.8&382\\ 
R22:S10&2.6&4.4&4.4&389\\
R22:S20&1.4&3.6&4.2&375\\
R22:S01&2.6&5.0&4.2&86\\
R22:S11&1.2&3.4&4.0&81\\
R22:S21&1.6&3.8&4.0&382\\
R22:S02&1.6&3.4&4.0&178\\
R22:S12&1.8&4.0&4.0&78\\
R22:S22&1.4&4.2&4.2&389\\
\hline
\end{tabular}
\caption{We looked at the resulting RMS from WCS fitting on the source lists of the deep exposure to the reference catalog, the science exposure to the reference catalog,
fitting the deep exposure directly to the science exposure, and fitting a warped coadd to the science exposure.  Fitting 
the coadd to the science exposure produces an order of magnitude more scatter in the fit than at any other point. {\bf SIMON MAKE SURE NUMBERS ARE STRAIGHT}\label{tab-wcsrms}}
\end{table}
\clearpage

>>>>

We found a significant dependence of the false positive rate on our
understanding of the variance; even a 2\% misestimate of the noise may
change the false positive rate by a factor of 1.5--2 at 5--sigma.
Prefiltering is preferred in this regard, as our variance planes are
shown to track the variance to higher fidelity at the detection stage.
We find that over/undersubtraction of the image background leads to a
bias in the positive to negative false detection ratio, but not on the
overall numbers of detections.  Prefiltering may also prevent a
deconvolution of the template image when the science image has better
seeing, which is a clear advantage of the technique.

@@

\subsection{Astrometric Registration \label{subsec-astrom}}
The original plan for this production was to use the single deep exposures (v6866600) as 
inputs to the coadd process, to exercise use of the coadd process and coadd repository.  As 
such we warped the exposures to coadd patches for use in the image difference
task.  The register task was unable to solve a differential WCS between the coadd
and science {\tt calexp} because in coadds, the parent WCS is for the entire {\tt tract}.
Each coadd {\tt patch} thus has a CRPIX at the center of the {\tt tract}, which is
typically off the {\tt patch} image.  TAN--SIP is unable to perform in this circumstance,
as it expands its polynomials around the CRPIX.  To get around this 
we tried re--solving the coadd patches using a TAN--SIP solution that was solved from
scratch.  This worked fairly will part of the time,
yielding residuals in the 40--100 mas range, compared to the 100--300 mas previously.  However,
as we did not understand the origin of these problems, and there were legitimate bugs
in meas\_astrom  that were uncovered as a result of this investigation, we reverted to
using the deep {\tt calexp} as the template directly; the direct deep {\tt calexp} to science {\tt calexp} 
registration typically yielded residuals in the range of 4--5 mas.

We investigated potential causes of the large residuals.  To do this, we looked at the 
residuals for template--to--reference catalog solutions, science image--to--refcat solutions, and 
direct template to science image registration to attempt to isolate where the astrometry 
was being degraded.  The results are shown
in Table~\ref{tab-wcsrms}.  We show the results for the nine sensors in the best seeing visit.  
The solutions for the three steps that do not involve any warping show results consistent with 
the accuracy we expect to get.  We also see the expected trend that the deep exposure solutions have 
smaller RMS than the solutions for the science images, due to their higher S/N.  The coadd--to--science 
registration depends critically on the WCS fitting and warping process: the WCS of the deep {\tt calexp} is 
warped into the coadd {\tt patch}; the coadd {\tt patch} is then warped into the science {\tt calexp's} WCS
frame.  This places two dependencies on the warping code, and two on the WCS fitting, where
this degradation may have occurred.  This 1--2 order of magnitude degradation in the WCS solutions 
shown in Table~\ref{tab-wcsrms} indicate
that astrometry and warping are areas that need to be actively addressed.

@@

\clearpage
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\multicolumn{5}{|c|}{WCS RMS in mas at various stages} \\
\hline
Sensor    & Deep & Science &  Deep to Science & Coadd to Science \\
\hline
R22:S00&4.2&5.8&3.8&382\\ 
R22:S10&2.6&4.4&4.4&389\\
R22:S20&1.4&3.6&4.2&375\\
R22:S01&2.6&5.0&4.2&86\\
R22:S11&1.2&3.4&4.0&81\\
R22:S21&1.6&3.8&4.0&382\\
R22:S02&1.6&3.4&4.0&178\\
R22:S12&1.8&4.0&4.0&78\\
R22:S22&1.4&4.2&4.2&389\\
\hline
\end{tabular}
\caption{We looked at the resulting RMS from WCS fitting on the source lists of the deep exposure to the reference catalog, the science exposure to the reference catalog,
fitting the deep exposure directly to the science exposure, and fitting a warped coadd to the science exposure.  Fitting 
the coadd to the science exposure produces 1--2 orders of magnitude more scatter in the fit than at any other point. \label{tab-wcsrms}}
\end{table}
\clearpage
